---
title: 'Assess fold changes based on the LFQ summaries'
output: html_notebook
author: Lieven Clement
---

We will use an R notebook to analyse the summarized data.
This will allow us to weave text and R commands.
The R chunks can be executed by pressing the play button in the top right above the R chunk.

#Data Import
We can download summarized data from the summarization tab of MSqRob. It is in Excel format.
Excel data can be read into R by the opelxlsx package.
We first load the package using the 'library(openxlsx)' function.
Object in R are assigned by the '=' or the '<-' operator.
The latter operator is generally preferred.
With the function head we can inspect the first few rows of a matrix or a dataframe.

##Load data
```{r}
library(openxlsx)
lfqSum<-read.xlsx("lfqSummariesLab3.xlsx")
head(lfqSum)
```

We will add a new column to the dataset that indicates if the peptide is a ups protein or a yeast protein.
We can do that by using the accession column. The peptides' name included the pattern "ups".
The "grepl" function in R can be used to check if a string contains a pattern. It returns a boolean: TRUE or FALSE.

##Extract UPS names
```{r}
#We first rename the
names(lfqSum)[1] <-"accession"
lfqSum$ups<-grepl(pattern="ups",lfqSum$accession)
head(lfqSum)
```

Next, we will calculate the mean for each group and the log2 fold change.
Finally we will make a boxplot for the ups and yeast proteins.
Note, that the real log2 fold change is 1.5. (Check this).

##Calculate means
```{r}
lfqSum$mu1<-rowMeans(lfqSum[,grepl(pattern="X6A",names(lfqSum))],na.rm=TRUE)
lfqSum$mu2<-rowMeans(lfqSum[,grepl(pattern="X6B",names(lfqSum))],na.rm=TRUE)
lfqSum$logfc<-lfqSum$mu2-lfqSum$mu1
head(lfqSum)
boxplot(logfc~ups,data=lfqSum,main=paste0(sum(!lfqSum$ups)," yeast and ",sum(lfqSum$ups)," ups proteins"))
abline(h=1.5,col="red",lwd=2)
```
